---
title: "Worksheet4c"
output: pdf_document
date: "2023-11-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg

1a. Show your solutions on how to import a csv file into the environment.

```{r}

mpg <- read.csv(file = "mpg.csv", header = T, sep = ",")
mpg

```

1b. Which variables from mpg dataset are categorical?
```{r}

```
1c. Which are continuous variables?
  
```{r}

```


2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.

```{r}


library(dplyr)

  mostmodels_manufacturer <- mpg %>%
  group_by(manufacturer) %>%
  summarise(num_models = n_distinct(model)) %>%
  arrange(desc(num_models)) %>%
  head(1)


  mostmodels_variations <- mpg %>%
  group_by(model) %>%
  summarise(num_variations = n()) %>%
  arrange(desc(num_variations)) %>%
  head(1)


cat("Manufacturer with the most models:", mostmodels_manufacturer $manufacturer, "\n")
cat("Model with the most variations:", mostmodels_variations$model, "\n")


```

2a. Group the manufacturers and find the unique models. Show your codes and result
```{r}

library(dplyr)

UniqueModels <- summarize(group_by(mpg, manufacturer), unique_models = unique(model))


print(UniqueModels)

```

2b. Graph the result by using plot() and ggplot(). Write the codes and its result.

```{r}

#Using plot

plot(Manufacturers_unique_models$Unique_models,
     names.arg = Manufacturers_unique_models$manufacturer,
     main = "Number of Unique Models by Manufacturer",
     xlab = "Manufacturer",
     ylab = "Number of Unique Models",
     col = "red",
     border = "black",
     las = 2,
     ylim = c(0, max(Manufacturers_unique_models$Unique_models) + 1))


library(ggplot2)

#using ggplot()

ggplot(Manufacturers_unique_models, aes(x = manufacturer, y = Unique_models)) +
  geom_bar(stat = "identity", fill = "pink", color = "purple") +
  labs(title = "Number of Unique Models by Manufacturer",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

2.1 Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.


2a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r} 
 
ggplot(mpg, aes(model, manufacturer)) + geom_point() 

#The "model" variable is represented by the x-axis. #The variable "manufacturer" is represented by the y-axis. 
#Every point on the plot represents a distinct pairing between a car model and its manufacturer.

#Points can be added to the plot using the geom_point() function. The points will be spread over the plot, with one point for each combination of model and manufacturer, as there are two categorical variables (model and manufacturer).

```
2b. For you, is it useful? If not, how could you modify the data to make it more informative?

```{r}




```
3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.
```{r}


```
4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}

```

4a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results.
```{r}

```















