---
title: "RWorksheet_AHUMADA#4b"
output: pdf_document
date: "2023-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix. 
```{r name, echo = TRUE}

vectorA <- c(1,2,3,4,5)

matrix_A <- matrix(0,nrow = 5, ncol =5)

for (j in 1:5)
  for (k in 1:5)
  {
    matrix_A[j,k] <- abs (vectorA[j] - vectorA[k])
  }

matrix_A

```
2. Print the string "*" using for() function. The output should be the same as shown in Figure1

```{r name, echo = TRUE}
for (j in 1:5) {
  cat(paste0("\"", rep("*", j), "\""), "\n")
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output. 

```{r name, echo = TRUE}

userinput <- as.integer(readline("Enter starting number for Fibonacci sequence: "))

x <- input
y <- 0


cat("Fibonacci sequence starting from", input, ":\n")

repeat {
 
  nextSeq <- x + y
 
  if (nextSeq > 500){
    break
  }
  cat(nextSeq, " ")
  x <- y
  y <- nextSeq
 
}

```
4. Import the dataset as shown in Figure 1 you have created previously.

4a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset?
Show your codes and its result.

```{r name, echo = TRUE}

write.csv(householdata, file = "householdata.csv")
imported<- read.csv("householdata.csv")
imported
importedx<-head(householdata, 6)
importedx


```
