---
title: "RWorksheet_AHUMADA#3b"
output: pdf_document
date: "2023-10-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r name, echo = TRUE} 
#1.a
resp_no <- c(1:20)
sex <- c(2,2,1,2,2,2,2,2,2,2,1,2,2,2,2,2,2,2,1,2)
occ <- c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1)
pers_at_home <- c(5,7,3,8,5,9,6,7,8,4,7,5,4,7,8,8,3,11,7,6)
sibs <- c(6,4,4,1,2,1,5,3,1,2,3,2,5,5,2,1,2,5,3,2)
h_type <- c(1,2,3,1,1,3,3,1,2,3,2,3,2,2,3,3,3,3,3,2)


household_data <- data.frame(
  Respondents = resp_no,
  Sex = sex,
  FatherOccupation = occ,
  PersonAtHome = pers_at_home,
  SiblingsAtSchool = sibs,
  HouseType = h_type
)
household_data

#1.2

str(household_data)
summary(household_data)

# 20 observations are represented by rows in the data frame, while 6 variables are represented by columns.

# the variables include:
# respondents - respondent's specific numeric identification
# sex - indicates the respondent's gender (1 for male, 2 for female).
# father's occupation - identifies the father's line of work (1 for farmer, 2 for driver, and 3 for others).
# persons at home - indicates the population of the house.
# siblings at school - reveals how many siblings are enrolled in school.
# type of house - gives a description of the type of home (1 for wood, 2 for semi-concrete, and 3 for concrete).

#1.c
sibs_mean <- mean(household_data$SiblingsAtSchool)

sibs_mean

# the mean of the number of siblings at school is 2.95, which is not 5

#1.d
firstTwoRows <- household_data[1:2,]
firstTwoRows

#1.e
thirdAndFifthRows <- household_data[c(3,5),c(2,4)]
thirdAndFifthRows


#1.f
types_houses <- household_data$HouseType
types_houses


#1.g
male_farmer <- household_data[household_data$Sex == 1 & household_data$FatherOccupation == 1,]

male_farmer

# there are no observations

#1.h

female_resp <- household_data[household_data$SiblingsAtSchool >= 5,]

female_resp

# there are five observations

# ------------------------------------

# 2
df = data.frame(Ints=integer(),
Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE
)

print("Structure of the empty dataframe:")
print(str(df))

# Created with 0 rows and 5 columns, the data frame df is empty.
# the columns contain the following type of data:
# ints = integer
# doubles = double
# characters = character
# logicals = logical
# factors  = factor (0 levels which means empty)

# acts as a template and may be filled with information

# -----------------------------------------------

# 3

new_resp <- c(1:10)
new_sex <- c("Male", "Female", "Female", "Male", "Male", "Female", "Female", "Male", "Female", "Male")
new_occ <- c(1,2,3,3,1,2,2,3,1,3)
new_personsAtHome <- c(5,7,3,8,6,4,4,2,11,6)
new_sibs <- c(2,3,0,5,2,3,1,2,6,2)
new_type <- c("Wood", "Congrete", "Congrete", "Wood", "Semi-congrete", "Semi-congrete", "Wood", "Semi-congrete", "Semi-congrete", "Congrete")

HouseholdData <- data.frame(
  Respondents = new_resp,
  Sex = new_sex,
  FatherOccupation = new_occ,
  PersonAtHome = new_personsAtHome,
  SiblingsAtSchool = new_sibs,
  HouseType = new_type
)

write.csv(HouseholdData, file = "HouseholdData.csv")

#3a

imported <-  read.csv("HouseholdData.csv")
imported

#3b

imported$Sex <- factor(imported$Sex, levels = c("Male", "Female"))
imported$Sex <- as.integer(imported$Sex)

imported$Sex

#3c

imported$HouseType <- factor(imported$HouseType, levels = c("Wood", "Congrete", "Semi-congrete"))
imported$HouseType <- as.integer(imported$HouseType)

imported$HouseType

#3d

imported$FatherOccupation <- factor(imported$FatherOccupation, levels = c(1,2,3), labels = c("Farmer", "Driver", "Others"))

imported$FatherOccupation

#3e

female_driver <- imported[imported$Sex == 2 & imported$FatherOccupation == "Driver",]
female_driver

#3f

greaterFive <- imported[imported$SiblingsAtSchool >= 5,]
greaterFive


#4

# We may infer from this data that public sentiment is sensitive to outside influences and that it changes over time.
